---
title: "NYC Flight Delay Report"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(tidyverse)
library(dplyr)
library(janitor)
library(DataExplorer)
library(nycflights13)
library(scales)
library(ggridges)
library(reactable)
flight <- nycflights13::flights %>%
    remove_constant() %>%
    drop_na(dep_time)

```

## Introduction and Sceniaro 

* Dataset introduction.
* Locations.
* Tables.

## Introduction and Scenarios
In our presentation we will be looking at the delay in:

* Different moths of the year
* Different airports
* The average delay for each carrier
* And the overall distribution of delays

## Exploratory Data Analysis
* The flights dataset has 328,521 rows and 19 columns.
* Three origin airport.
* 105 destination airports.
* 16 carriers.
* We removed the years column because it has the same value for all the rows, and we removed the rows that had missing values for the departure time.
 

## Exploratory Data Analysis
### Check which airport has the most flights
```{r}
flight %>%
    count(origin) %>%
    arrange(desc(n))
```

### Check which airport has the most delays
```{r}
flight %>%
    filter(dep_delay > 0) %>%
    count(origin) %>%
    arrange(desc(n))
```

## Exploratory Data Analysis
### Check the proportion of delay to the number flights

```{r}
flight %>%
    filter(dep_delay > 0) %>%
    count(origin) %>%
    mutate(prop = n / sum(n)) %>%
    arrange(desc(prop))
```

## Key investigations results from the data
### Which month has the most delays?

```{r}
flight %>%
  filter(dep_delay > 0) %>%
  group_by(month, origin) %>%
  summarise(count=n()) %>%
  ggplot(aes(month, count, color = origin))+
  geom_point(shape = 4) +
  geom_line() +
  scale_x_continuous("Month", n.breaks = 12) +
  scale_y_continuous("Count of delays") +
  ggtitle("Delays count by month from all NYC airport.") +
  theme_classic()
```

## Key investigations results from the data
### What is the distribution of all delays?

```{r}
ggplot(flight, aes(x = dep_delay)) +
  geom_histogram(bins = 80) +
  scale_x_continuous(limits = c(-100, 500))
```

## Key investigations results from the data
### What is the average delay for each airline?

```{r}
flight %>%
  group_by(carrier) %>%
  summarise(avg = mean(dep_delay), std = sd(dep_delay)) %>%
  ggplot(aes(carrier, avg)) +
  geom_col() +
  #geom_errorbar(aes(ymin=avg-std, ymax=avg+std))+
  scale_y_continuous("Average delay in min") +
  scale_x_discrete("Carrier") +
  ggtitle("Average delay in minute for each carrier thorughout 2013") +
  theme_classic()
```

## Takeaways

* Most flights and delays happen during the summer months.
* The proportion of delays increases with the number of flights.
* F9 has the highest average delay.


